---
title: "Homework 1"

author: "Maliha Safdar (ms7354)"

date: "2025-09-18"

output: github_document
---

```{r}
library(moderndive)
library(tidyverse)
data("early_january_weather", package = "moderndive")
```


#### **Problem 1: Solution**


```{r}
nrow(early_january_weather) # number of rows

ncol(early_january_weather) # number of columns
```

`In this first problem we will look at the behavior of the ggplot for different types of variables. The "early_january_weather" dataset has hourly meteorological data for LGA,JFK and EWR from January 2013. The size of the dataset is 358 rows and 15 columns. It has 15 variables and their names are:`

*   origin

*   year,month,day, hour, (for recording time)

*   temp, dewp (that are measured in F)

*   humid

*   wind_dir, wind_speed, wind_gust (wind direction in degrees) and (gust speed in mph)

*   precip (precipiitation, in inches)

*   visib (in miles)

*   time_hour (date and hour of the recording)`


```{r}
mean(early_january_weather$temp) # mean temperature
```

`The mean temperature is 39.58 F.`

##### Scatterplot for Temperature versus Time


```{r}
ggplot(early_january_weather, aes(x=time_hour, y=temp, color = humid)) +
  geom_point()

ggsave("scatter_plot.pdf", height = 4, width = 6)
```

`What I can observe from the scatterplot above is that as the temperature fluctuates between 40 degrees Farenheit to about 60 degrees Fareheit as time increases. The humidity can be seen to increase in the midpoint of Jan 7th and Jan 14th when the temperature is around 49 degrees Farenheit.`


#### **Problem 2: Solution**

`The sugar_df is a data frame that has information of amount of sugar in different fruits and the level of sugar each one has. It is a fictional dataframe I created so it is inaccurate.` 


```{r}

   vec_numeric = rnorm(10)
   vec_logical = vec_numeric > 0
   vec_char = c("apple","banana","cherry","dragonfruit","elderberry","fig","grapes","lychee","pineapple","mango")
     vec_factor = factor(c("high sugar level","high sugar level","medium sugar level","low sugar level","low sugar level","medium sugar level","medium sugar level","low sugar level","high sugar level","medium sugar level"))
```

```{r}
sugar_df <- data.frame(
     sugar_amount = vec_numeric,
     sugar_grt_ten = vec_logical,
     fruit = vec_char,
     sugar_level = vec_factor
     )
```


```{r}
(sugar_df)
```


```{r}
mean(sugar_df$sugar_amount) #mean for sugar_amount

mean(sugar_df$sugar_grt_ten) #mean for sugar_grt_ten

mean(sugar_df$fruit)        #mean for fruit

mean(sugar_df$sugar_level)  # mean for sugar_level
```

`When I took the mean of each variable I noticed that only the mean for sugar_amount (-0.38) and sugar_grt_0 (0.4) were populated. For the other two categorical variables no mean was populated.`

#### Using the pull function

`I pulled my variables from my sugar_df.`

```{r}
pull(sugar_df,sugar_amount)
pull(sugar_df,sugar_grt_ten)
pull(sugar_df,fruit)
pull(sugar_df,sugar_level)
```

#### Converting variables from one type to another using as.numeric
```{r, eval =FALSE}
as.numeric(sugar_df$sugar_grt_ten)
```
```{r, eval= FALSE}
as.numeric(sugar_df$fruit)
```
```{r, eval= FALSE}
as.numeric(sugar_df$sugar_level)
```

`When I used the as.numeric function to convert the logical, character and factor to numeric variable it did convert it for logical and factor but it doesn't make sense to take a mean for those values. For example, the logical variable when converted to numeric becomes 0 and 1 and the mean for that wouldn't be the same as what was computed for the numeric variable. Likewise, when the factor was converted to numeric it only produced the levels 1,2 and 3. There was no numeric conversion observed for character variable.`

