---
title: "P8105_HW1_ms7354"

author: "Maliha Safdar"

date: "2025-09-18"

output: github_document
---



#### **Problem 1: Solution**

```{r}
library(moderndive)
library(tidyverse)
data("early_january_weather", package = "moderndive")
```
```{r, eval=FALSE}
summary(early_january_weather)
```




```{r}
nrow(early_january_weather) # number of rows

ncol(early_january_weather) # number of columns
```

`The "early_january_weather" dataset has hourly meteorological data for LGA,JFK and EWR from January 2013. The size of the dataset is 358 rows and 15 columns. It has 15 variables and their names along with certain important values for some variables I identified are:`

*   origin

*   year,month,day, hour, (for recording time)

*   temp, dewp (that are measured in F)

*   humid, mean = 65.48

*   wind_dir (wind direction in degrees)

*   wind_speed (in mph), mean = 8.226 mph

*   wind_gust (in mph)

*   precip (precipitation, in inches), mean = 0.002039 in

*   pressure, mean = 1023

*   visib (in miles)

*   time_hour (date and hour of the recording)`


```{r}
mean(early_january_weather$temp) # mean temperature
```

`The mean temperature is 39.58 F.`

##### Scatterplot for Temperature versus Time


```{r}
ggplot(early_january_weather, aes(x=time_hour, y=temp, color = humid)) +
  geom_point()

ggsave("scatter_plot.pdf", height = 4, width = 6)
```

`What I can observe from the scatterplot above is that as the temperature fluctuates between 40 degrees Farenheit to about 60 degrees Fareheit as time increases. The humidity can be seen to increase in the midpoint of Jan 7th and Jan 14th when the temperature is around 49 degrees Farenheit.`


#### **Problem 2: Solution**

`The sugar_df is a data frame that has information of amount of sugar in different fruits and the level of sugar each one has. It is a fictional dataframe I created so it is inaccurate.` 


```{r}
set.seed(1234)
   vec_numeric = rnorm(10)
   vec_logical = vec_numeric > 0
   vec_char = c("apple","banana","cherry","dragonfruit","elderberry","fig","grapes","lychee","pineapple","mango")
     vec_factor = factor(c("high sugar level","high sugar level","medium sugar level","low sugar level","low sugar level","medium sugar level","medium sugar level","low sugar level","high sugar level","medium sugar level"))
```

```{r}
sugar_df <- data.frame(
     sugar_amount = vec_numeric,
     sugar_grt_ten = vec_logical,
     fruit = vec_char,
     sugar_level = vec_factor
     )
```


```{r}
(sugar_df)
```
##### Calculating mean of each variable

```{r}
mean(sugar_df$sugar_amount) #mean for sugar_amount

mean(sugar_df$sugar_grt_ten) #mean for sugar_grt_ten

mean(sugar_df$fruit)        #mean for fruit

mean(sugar_df$sugar_level)  # mean for sugar_level
```

`When I took the mean of each variable I noticed that only the mean for my numeric variable, sugar_amount (-0.38) logical variable, sugar_grt_0 (0.4) were populated. For the other two categorical variables no mean was populated. Furthermore, the mean for the logical variable is just the sum of 1 and 0 (which are values for "TRUE" and "FALSE") which does not really reprsent the mean.`

#### Using the pull function

`The tidyverse has already been loaded so I didn't load it again. I pulled my variables from my sugar_df.`

```{r}
pull(sugar_df,sugar_amount)
pull(sugar_df,sugar_grt_ten)
pull(sugar_df,fruit)
pull(sugar_df,sugar_level)
```

#### Converting variables from one type to another using as.numeric
```{r, eval =FALSE}
as.numeric(sugar_df$sugar_grt_ten)
```
```{r, eval= FALSE}
as.numeric(sugar_df$fruit)
```
```{r, eval= FALSE}
as.numeric(sugar_df$sugar_level)
```

`When I used the as.numeric function to convert the logical, character and factor to numeric variable it did convert to numeric variables for logical and factor but it doesn't make sense to take a mean for those values. For example, the logical variable when converted to numeric becomes 0 and 1 and the mean for that wouldn't be the same as what was computed for the numeric variable. Likewise, when the factor was converted to numeric it only produced the levels 1,2 and 3. There was no numeric conversion observed for character variable.`

